Paper 2\thanks{Code and data are available at: LINK.https://github.com/Crooksyyy/The-Effects-of-Social-Media , Original data available https://www.openicpsr.org/openicpsr/project/112081/version/V1/view}

### 2024-02-15 outputs/paper/paper.qmd
almost done

---
title: "Paper 2"
author: 
  - Gavin Crooks
  - Samarth Rajani
thanks: "Code and data are available at: LINK.https://github.com/Crooksyyy/The-Effects-of-Social-Media"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
toc: true
bibliography: references.bib 
nocite: |
  @*
---

@@

# Introduction
Household income is defined as the gross income earned by all members in a household above 15 years of age (https://www.investopedia.com/terms/h/household_income.asp). Over the years, it has been debated whether household incomes at all affect one’s affiliation toward a political school of thought. It is a reasonable hypothesis to assume a sort of relationship between income and voting either Democrat or Republican, as both parties have different economic outlooks thereby affecting incomes differently. Presumably, a higher income inequality polarizes political leaning further. Therefore it is in our best interests to study whether the economically disadvantaged vote to improve their quality of life.

In ‘Income Inequality and Partisan Voting in the United States’, Andrew Gelman, Lane Kenworthy and Yu-Sung Su (https://www.jstor.org/stable/42956457?seq=11) make a case for higher earning Americans voting Republican, whereas Jeff Madrick (https://www.nytimes.com/2020/07/31/books/review/the-system-robert-reich-break-em-up-zephyr-teachout.html) argues how working-class Americans voted supposedly against their interests in voting Republican. Conflicting theories have emerged, and we intend to tackle this issue at hand of whether different income brackets tend to vote differently.

The introduction is self-contained and tells a reader everything they need to know including: 1) broader context to motivate; 2) some detail about what the paper is about; 3) a clear gap that needs to be filled; 4) what was done; 5) what was found; 6) why it is important; 7) the structure of the paper. A reader should be able to read only the introduction and know what was done, why, and what was found. Likely 3 or 4 paragraphs, or 10 per cent of total.

@@

The third variable in the data that is included in our analysis is a variable of respondents self identifying how closely they follow politics. This is another categorical variable measured as Not at all closely, Somewhat closely, Rather closely and  Very closely. This variable faces many problems as this categorical scale is not consistent across respondents. To be specific we mean someone who identifies as someone who does follows Not at all closely can be following politics more than someone who identifies as Somewhat closely. This is a measurement issue within to the questions asked in the survey and all self identifying variables in general. @figure2 shows the quantity of respondents in each group. The most common response is that they follow somewhat closely and the other responses are relatively even. 
```{r}
#| label: fig-figure2
#| fig-cap: "How Closely People Follow Politics from Responses in a facebook ad"
#| warning: false
#| message: false
#| echo: false

ggplot(cleaned_data, aes(x = "", fill = follow_politics)) +
  geom_bar() +
  coord_polar("y") +
  labs(fill = "Legend") +
  theme(axis.title = element_blank())

@@

```{r}
# Parts of this code from R-charts
#https://r-charts.com/part-whole/stacked-bar-chart-ggplot2/
# Assisited by Chat GPT3.5 - view full chat in inputs/llms/usage.txt
ggplot(cleaned_data, aes(x = follow_trump, fill = race)) +
  geom_bar(position = 'stack', color = 'black', stat = 'count') +
  labs(title = "Number of People Following Trump by Race",
       x = "Follows Trump",
       y = "Count") +
  scale_fill_manual(values = c('White / Caucasian' = 'lightblue', 'Black or African American' = 'lightcoral', 'Asian or Pacific Islander' = 'lightgreen', 'Hispanic' = 'plum',
                               'Other  (please specify)' = 'lightgoldenrodyellow')) +
  theme_minimal()

@@

```
# Model

simple regression of income and politics data controlling for race

>>>>

---
title: "Paper 2"
author: 
  - Gavin Crooks
  - Samarth Rajani
thanks: "Code and data are available at: LINK.https://github.com/Crooksyyy/The-Effects-of-Social-Media , Original data available https://www.openicpsr.org/openicpsr/project/112081/version/V1/view"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
toc: true
bibliography: references.bib 
nocite: |
  @*
---

@@

# Introduction {#intro}
Household income is defined as the gross income earned by all members in a household above 15 years of age [@hhld]. Over the years, it has been debated whether household incomes at all affect one’s affiliation towards a political school of thought. It is a reasonable hypothesis to assume a sort of relationship between income and voting either Democrat or Republican, as both parties have different economic outlooks thereby affecting incomes differently. Maybe higher income inequality polarizes political leaning further. Therefore, it is in our best interests to study whether the poor vote to improve their quality of life.

In ‘Income Inequality and Partisan Voting in the United States’, Andrew Gelman, Lane Kenworthy and Yu-Sung Su [@lane] make a case for higher earning Americans voting Republican, whereas Jeff Madrick [@article] argues how working-class Americans voted against their interests in voting Republican. Conflicting theories have emerged, and we intend on tackling this issue at hand of whether different income brackets tend to vote differently.

Our data has been obtained from ‘The Welfare Effects of Social Media’ [@paper] .
Our code is supported by the following packages [@citeR] [@tidy] [@her] [@knitr]

@@

The third variable in the data that is included in our analysis is a variable of respondents self identifying how closely they follow politics. This is another categorical variable measured as Not at all closely, Somewhat closely, Rather closely and  Very closely. This variable faces many problems as this categorical scale is not consistent across respondents. To be specific we mean someone who identifies as someone who does follows Not at all closely can be following politics more than someone who identifies as Somewhat closely. This is a measurement issue within to the questions asked in the survey and all self identifying variables in general. @figure2 shows the quantity of respondents in each group. The most common response is that they follow somewhat closely and the other responses are relatively even. 
```{r}
#| label: fig-figure2
#| fig-cap: "How Closely People Follow Mr.Donald Trump from Responses in a facebook ad"
#| warning: false
#| message: false
#| echo: false

ggplot(cleaned_data, aes(x = "", fill = follow_trump)) +
  geom_bar() +
  coord_polar("y") +
  labs(fill = "Legend") +
  theme(axis.title = element_blank())

@@

```{r}
#| label: fig-figure4
#| fig-cap: "Number of Respondents who follow Donald Trump at different levels by Race"
#| warning: false
#| message: false
#| echo: false
# Parts of this code from R-charts
#https://r-charts.com/part-whole/stacked-bar-chart-ggplot2/
# Assisted by Chat GPT3.5 - view full chat in inputs/llms/usage.txt
ggplot(cleaned_data, aes(x = follow_trump, fill = race)) +
  geom_bar(position = 'stack', color = 'black', stat = 'count') +
  labs(title = "Number of People Following Trump by Race",
       x = "Follows Trump",
       y = "Count") +
  scale_fill_manual(values = c('White / Caucasian' = 'lightblue', 'Black or African American' = 'lightcoral', 'Asian or Pacific Islander' = 'lightgreen', 'Hispanic' = 'plum',
                               'Other  (please specify)' = 'lightgoldenrodyellow')) +
  theme_minimal()

@@

```
# Results

```{r}
# Create a table of proportions
income_follow_table <- table(cleaned_data$follow_trump, cleaned_data$income_group)
income_proportions <- prop.table(income_follow_table, margin = 1) * 100

table1 <- tibble(income_proportions, colnames(c("Income Group", "Follows Trump" )))
show(table1)
```

This papers goal was to identify if lower income household voted against their own interest. To understand this relationship we used the variable for how closely a respondent follows Mr. Donald Trump as our measurement of subscription to Republican ideas. Using this measurement, we organized the proportion of individuals by income class to how closely they follow Mr. Donald Trump in @fig-figure3. This graph shows that a proportional amount of each income class follows Mr. Donald Trump at similar levels across all income levels. Specifically, we mean the percentage of individuals  follow Mr. Donald Trump at different levels is the same no matter the income class. This means we can not conclude that income class impacts how closely individuals follow Mr. Donald Trump, and therefor, cannot conclude that different income levels subscribe to republican ideas more than the other.

To further our analysis we computed the same graph however organized by race not income @fig-figure4. This resulted in a similar result as race is proportional between all levels of following Mr. Donald Trump. Therefore, consistent across races at to following republican rhetoric. Again, this means the same percentage of people that follow Mr. Donald Trump at different levels is the same for each race. Obviously, this is more difficult to conclude for minorities as their representation within the data set is so small as stated in @sec-race_data

Overall, the results of this analysis were inconclusive to measure how income impacts individuals propensity to follow republican rhetoric. There are many reasons this could be true and as stated in the @intro there are multiple schools of thought previously studies on the topic.

### 2024-02-15 outputs/paper/paper.tex
update again
almost done

\title{Paper 2\thanks{Code and data are available at:
LINK.https://github.com/Crooksyyy/The-Effects-of-Social-Media}}
\author{Gavin Crooks \and Samarth Rajani}
\date{February 14, 2024}

@@

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\section{Introduction}\label{introduction}

The introduction is self-contained and tells a reader everything they
need to know including: 1) broader context to motivate; 2) some detail
about what the paper is about; 3) a clear gap that needs to be filled;
4) what was done; 5) what was found; 6) why it is important; 7) the
structure of the paper. A reader should be able to read only the
introduction and know what was done, why, and what was found. Likely 3
or 4 paragraphs, or 10 per cent of total.

@@

\caption{\label{fig-figure2}How Closely People Follow Politics from
Responses in a facebook ad}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Parts of this code from R{-}charts}
\CommentTok{\#https://r{-}charts.com/part{-}whole/stacked{-}bar{-}chart{-}ggplot2/}
\FunctionTok{ggplot}\NormalTok{(cleaned\_data, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ follow\_trump, }\AttributeTok{fill =}\NormalTok{ race)) }\SpecialCharTok{+}
  \FunctionTok{geom\_bar}\NormalTok{(}\AttributeTok{position =} \StringTok{\textquotesingle{}stack\textquotesingle{}}\NormalTok{, }\AttributeTok{color =} \StringTok{\textquotesingle{}black\textquotesingle{}}\NormalTok{, }\AttributeTok{stat =} \StringTok{\textquotesingle{}count\textquotesingle{}}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{title =} \StringTok{"Number of People Following Trump by Race"}\NormalTok{,}
       \AttributeTok{x =} \StringTok{"Follows Trump"}\NormalTok{,}
       \AttributeTok{y =} \StringTok{"Count"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_fill\_manual}\NormalTok{(}\AttributeTok{values =} \FunctionTok{c}\NormalTok{(}\StringTok{\textquotesingle{}White / Caucasian\textquotesingle{}} \OtherTok{=} \StringTok{\textquotesingle{}lightblue\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}Black or African American\textquotesingle{}} \OtherTok{=} \StringTok{\textquotesingle{}lightcoral\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}Asian or Pacific Islander\textquotesingle{}} \OtherTok{=} \StringTok{\textquotesingle{}lightgreen\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}Hispanic\textquotesingle{}} \OtherTok{=} \StringTok{\textquotesingle{}plum\textquotesingle{}}\NormalTok{,}
                               \StringTok{\textquotesingle{}Other  (please specify)\textquotesingle{}} \OtherTok{=} \StringTok{\textquotesingle{}lightgoldenrodyellow\textquotesingle{}}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-pdf/unnamed-chunk-5-1.pdf}

\section{Model}\label{model}

simple regression of income and politics data controlling for race

@@

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-citeLahman}
Friendly, Michael, Chris Dalzell, Martin Monkman, and Dennis Murphy.
2020. \emph{Lahman: Sean {``Lahman''} Baseball Database}.
\url{https://CRAN.R-project.org/package=Lahman}.

\bibitem[\citeproctext]{ref-gebru2021datasheets}
Gebru, Timnit, Jamie Morgenstern, Briana Vecchione, Jennifer Wortman
Vaughan, Hanna Wallach, Hal Daumé Iii, and Kate Crawford. 2021.
{``Datasheets for Datasets.''} \emph{Communications of the ACM} 64 (12):
86--92.

\bibitem[\citeproctext]{ref-rstanarm}
Goodrich, Ben, Jonah Gabry, Imad Ali, and Sam Brilleman. 2022.
{``Rstanarm: {Bayesian} Applied Regression Modeling via {Stan}.''}
\url{https://mc-stan.org/rstanarm/}.

\bibitem[\citeproctext]{ref-palmerpenguins}
Horst, Allison Marie, Alison Presmanes Hill, and Kristen B Gorman. 2020.
\emph{Palmerpenguins: Palmer Archipelago (Antarctica) Penguin Data}.
\url{https://doi.org/10.5281/zenodo.3960218}.

@@

\bibitem[\citeproctext]{ref-thereferencecanbewhatever}
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

>>>>

\title{Paper 2\thanks{Code and data are available at:
LINK.https://github.com/Crooksyyy/The-Effects-of-Social-Media , Original
data available
https://www.openicpsr.org/openicpsr/project/112081/version/V1/view}}
\author{Gavin Crooks \and Samarth Rajani}
\date{February 15, 2024}

@@

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\section{Introduction}\label{intro}

Household income is defined as the gross income earned by all members in
a household above 15 years of age (SCOTT 2024). Over the years, it has
been debated whether household incomes at all affect one's affiliation
towards a political school of thought. It is a reasonable hypothesis to
assume a sort of relationship between income and voting either Democrat
or Republican, as both parties have different economic outlooks thereby
affecting incomes differently. Maybe higher income inequality polarizes
political leaning further. Therefore, it is in our best interests to
study whether the poor vote to improve their quality of life.

In `Income Inequality and Partisan Voting in the United States', Andrew
Gelman, Lane Kenworthy and Yu-Sung Su (Gelman, Kenworthy, and Su 2010)
make a case for higher earning Americans voting Republican, whereas Jeff
Madrick (\textbf{article?}) argues how working-class Americans voted
against their interests in voting Republican. Conflicting theories have
emerged, and we intend on tackling this issue at hand of whether
different income brackets tend to vote differently.

Our data has been obtained from `The Welfare Effects of Social Media'
(\textbf{paper?}) . Our code is supported by the following packages (R
Core Team 2022) (Wickham et al. 2019) (\textbf{her?}) (Xie 2023)

@@

\caption{\label{fig-figure2}How Closely People Follow Mr.Donald Trump
from Responses in a facebook ad}

@@

\includegraphics{paper_files/figure-pdf/fig-figure4-1.pdf}

\caption{\label{fig-figure4}Number of Respondents who follow Donald
Trump at different levels by Race}

\includegraphics{paper_files/figure-pdf/fig-figure3-1.pdf}

\caption{\label{fig-figure3}Number of Respondents who follow Donald
Trump at different levels by Household Income}

\section{Results}\label{results}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Create a table of proportions}
\NormalTok{income\_follow\_table }\OtherTok{\textless{}{-}} \FunctionTok{table}\NormalTok{(cleaned\_data}\SpecialCharTok{$}\NormalTok{follow\_trump, cleaned\_data}\SpecialCharTok{$}\NormalTok{income\_group)}
\NormalTok{income\_proportions }\OtherTok{\textless{}{-}} \FunctionTok{prop.table}\NormalTok{(income\_follow\_table, }\AttributeTok{margin =} \DecValTok{1}\NormalTok{) }\SpecialCharTok{*} \DecValTok{100}

\NormalTok{table1 }\OtherTok{\textless{}{-}} \FunctionTok{tibble}\NormalTok{(income\_proportions, }\FunctionTok{colnames}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"Income Group"}\NormalTok{, }\StringTok{"Follows Trump"}\NormalTok{ )))}
\FunctionTok{show}\NormalTok{(table1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 4 x 1
  income_proportions[,"$0 to $19,9~1 [,"$100,000 and up"] [,"$20,000 to $49,99~2
                               <dbl>                <dbl>                  <dbl>
1                               9.12                 16.9                   28.1
2                               5.86                 22.5                   29.0
3                               7.57                 20.2                   29.3
4                               6.73                 22.4                   25.3
# i abbreviated names: 1: income_proportions[,"$0 to $19,999"],
#   2: [,"$20,000 to $49,999"]
# i 1 more variable: income_proportions[4:5] <dbl>
\end{verbatim}

This papers goal was to identify if lower income household voted against
their own interest. To understand this relationship we used the variable
for how closely a respondent follows Mr.~Donald Trump as our measurement
of subscription to Republican ideas. Using this measurement, we
organized the proportion of individuals by income class to how closely
they follow Mr.~Donald Trump in Figure~\ref{fig-figure3}. This graph
shows that a proportional amount of each income class follows Mr.~Donald
Trump at similar levels across all income levels. Specifically, we mean
the percentage of individuals follow Mr.~Donald Trump at different
levels is the same no matter the income class. This means we can not
conclude that income class impacts how closely individuals follow
Mr.~Donald Trump, and therefor, cannot conclude that different income
levels subscribe to republican ideas more than the other.

To further our analysis we computed the same graph however organized by
race not income Figure~\ref{fig-figure4}. This resulted in a similar
result as race is proportional between all levels of following
Mr.~Donald Trump. Therefore, consistent across races at to following
republican rhetoric. Again, this means the same percentage of people
that follow Mr.~Donald Trump at different levels is the same for each
race. Obviously, this is more difficult to conclude for minorities as
their representation within the data set is so small as stated in
Section~\ref{sec-race_data}

Overall, the results of this analysis were inconclusive to measure how
income impacts individuals propensity to follow republican rhetoric.
There are many reasons this could be true and as stated in the
(\textbf{intro?}) there are multiple schools of thought previously
studies on the topic.

@@

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-10.1257ux2faer.20190658}
Allcott, Hunt, Luca Braghieri, Sarah Eichmeyer, and Matthew Gentzkow.
2020. {``The Welfare Effects of Social Media.''} \emph{American Economic
Review}. \url{https://doi.org/10.1257/aer.20190658}.

\bibitem[\citeproctext]{ref-lane}
Gelman, Andrew, Lane Kenworthy, and Yu-Sung Su. 2010. {``Income
Inequality and Partisan Voting in the United States.''} \emph{Social
Science Quarterly}. {[}University of Texas Press, Wiley{]}.
\url{http://www.jstor.org/stable/42956457}.

\bibitem[\citeproctext]{ref-Madrick_2020}
Madrick, Jeff. 2020. {``Why the Working Class Votes Against Its Economic
Interests.''} \emph{The New York Times}, July.
\url{https://www.nytimes.com/2020/07/31/books/review/the-system-robert-reich-break-em-up-zephyr-teachout.html}.

\bibitem[\citeproctext]{ref-here}
Müller, Kirill. 2020. {``Here: A Simpler Way to Find Your Files.''}
\url{https://CRAN.R-project.org/package=here}.

@@

\bibitem[\citeproctext]{ref-hhld}
SCOTT, MICHELLE P. 2024. {``Household Income.''} investopedia.
\url{https://www.investopedia.com/terms/h/household_income.asp}.

\bibitem[\citeproctext]{ref-tidy}
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source
Software}. \url{https://doi.org/10.21105/joss.01686}.

\bibitem[\citeproctext]{ref-knitr}
Xie, Yihui. 2023. {``Knitr: A General-Purpose Package for Dynamic Report
Generation in r.''} \url{https://yihui.org/knitr/}.
